Discrete Probability for Discrete Quantum Theories

### 2016-05-07 prob.tex
buffon

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware computational or experimental setting,
it is more appropriate to consider probability measures that map
events to a finite set of elements computable with a fixed set of
resources. We will consider two approaches from the literature:
set-valued probability measures~\cite{Artstein1972,PuriRalescu1983}
and interval-valued probability
measures~\cite{Dempster1967,Zadeh1986,Weichselberger2000,JamisonLodwick2004}.

>>>>

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware computational or experimental setting,
it is more appropriate to consider probability measures that map
events to a finite set of elements computable with a fixed set of
resources. We expand on this observation and then consider two
approaches from the literature: set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983} and interval-valued
probability
measures~\cite{Dempster1967,Zadeh1986,Weichselberger2000,JamisonLodwick2004}.

%%%%%
\subsection{Buffon's Needle Problem}

Suppose we drop a needle of length $\ell$ onto a floor made of equally
spaced parallel lines a distance $\ell$ apart. It is a known fact that
the probability of the needle crossing a line is
$\frac{2}{\pi} = 0.6366197723675814$. We analyze this situation in the
mathematical framework of probability spaces paying special attention
to the resources needed to estimate the probability computationally or
experimentally.

To formalize the experiment, we consider an experimental setup
consisting of a collection of $N$ identical needles of length
$\ell$. We throw the $N$ needles and observe the number $X$ of needles
that cross a line. The sample space can be expressed as the set
$(X|-)^N$ of sequences of characters of length $N$ where each
character is either $X$ to indicate a needle crossing a line or $-$ to
indicate a needle not crossing a line. If $N=3$, the event that
exactly 2 needles cross lines is $\{-XX,X-X,XX-\}$: it has
probability~$\frac{2}{3}$. Generally, the event that exactly $M$
needles out of the $N$ total needles cross lines is $\frac{M}{N}$.

In an actual experiment with 500 needles, it was found that 321
crossed a line which gives a probability of $0.642$. In a larger
experiment with $3408$ needles, the probability was
$0.6366197193098592$. Comparing the values to the idealized
mathematical value, we see that the observed probability approaches
$\frac{2}{\pi}$ but only if \emph{larger and larger resources} are
expended.

These resource considerations thus lead us to replace the real
interval $[0,1]$ with intervals of confidence which contain the
probability. The intuition is that the precision of the interval will
depend on the amount of resources allocated for the measurement.
